---
title: "ESM 204 - Assignment 3"
author: "Conner Smith"
date: "5/2/2022"
output:
  html_document:
   theme:
     bg: "#002B36"
     fg: "#EEE8D5"
     primary: "#2AA198"
   code_folding: hide
   
---

```{r setup, include=TRUE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, options(scipen = 999))

library(tidyverse)
library(here)
library(janitor)
library(thematic)
library(scales)
library(equatiomatic)

thematic::thematic_rmd()
thematic::thematic_on()

```


## Overview

```{r}
# Read in the main data set

carbon <- read_csv(here("carbon.csv")) %>% 
  clean_names() %>% 
  select(-x1)

```


The Biden Administration’s “interim” value is $51 per metric ton of CO2. The electricity sector is the second largest source of greenhouse gas emissions in the U.S. (after transportation). In this analysis, we consider the distributional consequences of imposing a household electricity tax based on the SCC to address the climate change problem.


## **Analysis** {.tabset}

The dataset contains the following:

- Consumers can separated into two income groups: “high” and “low.” 
- Price (in $) and quantity (in kWh) estimates of demand per month for the two groups. 

This analysis will start with linear regressions (with an intercept) to estimate the demand curves for “high” and “low” income consumers under the following conditions: 

- Initially, there is no tax on electricity consumption.
- The current electricity price (without any taxes) is $.10 per kWh.
- The marginal cost of producing a kWh of electricity is linear and has a price-intercept of 0.


### **1. MEC**

Assuming that one kWh of electricity emits 0.85 pounds of CO2 and that the interim SCC of $51/ton is accurate, this section calculates the MEC per kWh.

```{r}
# Create a longer dataset where income is a variable, not a column 

carbon_long <- carbon %>% 
  pivot_longer(cols = c(q_low_kwh, q_high_kwh),
               names_to = 'income_level',
               values_to = 'kwh') %>% 
  mutate(income_level = case_when(income_level == 'q_low_kwh' ~ 'low',
                   income_level == 'q_high_kwh' ~ 'high'))

# Run a simple calculation to get the MEC 

interim_price_ton <- 51

#2205 lbs in a metric ton. Divide by 0.85 to get kwh produced per ton

kwh_ton <- 2204.6/0.85

# Multiply by Biden price to get MEC/kWh (Basically, $51 is the price for 2594 kwh)
interim_price_kwh <- 51/kwh_ton

```

Based on this, the marginal externality cost per kWh of electricity is $`r round(interim_price_kwh, 4)`.

### **2. Aggregations**

**Low income demand curve:**
```{r}
# Calculate aggregate demand curve for electricity

# First get lm estimates from data for both low and high
demand_low <- lm(price_cents ~ kwh, income_level=='low', 
                 data = carbon_long) 
extract_eq(model = demand_low, use_coefs = TRUE, coef_digits = 5)

demand_low_f <- function(p){
  q <- (p - demand_low$coefficients[[1]])/
  demand_low$coefficients[[2]]
  return(q)}
```

**High income demand curve:**
```{r}
demand_high <- lm(price_cents ~ kwh, income_level=='high',
                  data = carbon_long) 
extract_eq(model = demand_high, use_coefs = TRUE, coef_digits = 5)

demand_high_f <- function(p){
  q <- (p - demand_high$coefficients[[1]])/
  demand_high$coefficients[[2]]
  return(q)}

```

Given these income-based demand curves, we can sum horizontally to get the **aggregate demand curve:**

```{r}
# Switch the equations above so Q is on one side and then add 

demand_low_flip <- lm(kwh ~ price_cents, income_level=='low', 
                 data = carbon_long) 

demand_high_flip <- lm(kwh ~ price_cents, income_level=='high',
                  data = carbon_long) 

kwh_agg_int <- (demand_low_flip$coefficients[1]+
                  demand_high_flip$coefficients[1])

kwh_agg_slope <- (demand_low_flip$coefficients[2] +
                    demand_high_flip$coefficients[2])

demand_agg_p <- function(price_cents){kwh_agg_int +
    kwh_agg_slope*(price_cents)}

demand_agg_int <- kwh_agg_int/-kwh_agg_slope
demand_agg_slope <- 1/kwh_agg_slope

# Aggregate demand function is P = 30.5 - 0.000039Q

demand_agg <- function(kwh){demand_agg_int + demand_agg_slope*kwh}

demand <- function(p, model){
  q <- (p - model$coefficients[[1]])/model$coefficients[[2]]
  q <- ifelse(q<0,0,q)
  return(q)
}

```

This gives us an aggregate demand of equation of *P = 30.5 - 0.000039Q*. 

*Note: this gives price in cents, not dollars*

Now we need to calculate the supply curve for electricity. We already know that the current price is $0.10/kWh. We also know the MC for electricity is linear with a 0 intercept. 

```{r}

kwh_agg <- demand_agg_p(10)
# 522128 kWh consumed at price of $0.10

supply_slope <- 10/kwh_agg
```

Since we know the supply curve will pass through the aggregate demand curve at the level of demand for a price of $0.10/kwh, we can infer that the supply curve (or marginal cost per kWh of electricity) function is: 

*P = 0.0000192Q*

*Note: this gives a price in cents, not dollars*

From this, we can calculate consumer and producer surplus as well as environmental cost. 

```{r}
# CS needs Y intercept --> at 0 kWh, price would be $0.305

# Note the conversion back to dollars 
cs_baseline <- 0.5*kwh_agg*(0.305 - 0.10)
# $53,518

# PS easy, just go up to 10 and out to the kwh 
ps_baseline <- 0.5*kwh_agg*0.10
# $26,106

# cost is rectangle 
env_cost_baseline <- interim_price_kwh*kwh_agg
# $10,267
```

- Consumer Benefit = $`r round(cs_baseline, 0)`
- Producer Benefit = $`r round(ps_baseline, 0)`
- Environmental Cost = $`r round(env_cost_baseline, 0)`

*Note these benefits and costs are in dollars, not cents*

### **3. Consumer Benefit**
 
From above, we have a total consumer surplus of $53,518.
This comes from an aggregate demand curve that indicates 522,128 kWh will be consumed at  price of $0.10/kWh. 

To get the surplus differential between income groups, we use the current price and calculate the amount each group will consume at this level. 

```{r}
# low income demand function
demand_low_f(10)
#Low income gets 121,344 
#demand_low$coefficients[1] --- intercepot is 23.37
cs_low <- 0.5*(.2337 - .10)*demand_low_f(10)
# $8,112 CS for low income (note dollars/cents conversion)

demand_high_f(10)
#High gets 415,376
#demand_high$coefficients[1] --- intercept is 31.61
cs_high <- 0.5*(.3161-.10)*demand_high_f(10)
# $44,881.3 CS for high income 
```

This indicates clear disparity where consumer surplus for high income consumers is much higher than that for low income consumers. The difference is as follows: 

- Low-income CS: $`r round(cs_low, 0)`
- High-income CS: $`r round(cs_high, 0)`

### **4. Optimal Tax**

### **5. Redistribution**

### **6. Solar PV**